---
title: "Homework 02"
author: "Gustavo Baroni"
date: "3/20/2020"
output: pdf_document
---
## Basic R Operations  
## 1.1

R can help me to find how many seconds there are in a decade, for instance:  
There are 3600 seconds in an hour:  

```{r}
60*60
```

And, there are 86400 seconds in a day:   
```{r}
3600*24  
```

OK, we already know how many seconds exist in a day. Now, let's see how many days there are in a decade, considering two leap years:

```{r}
(366*2) + (365*8) 
```

There are 3652 days in a decade.  

Thus, I just nedd to calculate the quantity of days times the quantity of seconds in a day.  

```{r}
3652*86400 
```

Hence, there are 315532800 seconds in a decade. 

## 1.2
To install devtools, I can use install.packages:  
```
install.packages("devtools")
```

## 1.3
Installing ggplot2:  
```{r}
library(devtools)
install_github("tidyverse/ggplot2")
```

## 1.4
In order to learn more about separate() function:  
```{r}
library(tidyverse)
library(tidyr)
help("separate")
```

## 1.5
Making a pie chart:  
```{r}
x <- c(32, 21, 13, 34)
labels <- c("burger", "italian", "soup", "pizza")
pie(x, labels, main = "Dishes on Friday night")
```

## 1.6  
There are:  
```{r}
apropos("test")
```

## 1.7 
I've searched "Zelig" on StackOverflow and somebody had this problem: he or she could run something (named probitgee) using "Zelig". However, he or she could not get the output in a "stargazer table". The answer, actually, was pretty simple: someone said that it was necessary access the zelig.out model object to get the output. All the discussion is in https://stackoverflow.com/questions/49187185/stargazer-table-from-zelig-probitgee  

## 1.8 
I would use the "install.package" and I would use "data" to load this package.  

## 1.9
In R, I could use "help" to learn how to use the function. Out of R, I could just google it and, if I had not found a good answer, I could also search in StackOverflow.  

## 1.10
When I make a vector using only c(), it generates "NULL":  
```{r}
y <- c()
```

## 1.11
The generated vector is "":  
```{r}
e <- c("")
```

## 1.12
The single vector has only a class and in this case is "character":  
```{r}
x <- c(1,2,3)
y <- c('1','2','3')
c(x,y)
```

## 1.13
The variable cylinders shows the quantity of cylinders in the cars (32 cars in all):
```{r}
data("mtcars")
```

## 1.14
I am going to make a table of the variable cylinders:  
```{r}
table(mtcars$cyl)
```

## 1.15
All three show examples of misuse of data and graphs. There are, perhaps, two explanations to answer why that graphs are wrong: the first one is that who make the graph wanted to fool the readers and the second is that who make the graph did not know how to do properly.  

## 1.16 

 The chance that it will rain tomorrow: continuous quantitative.  
 Numbers of cars that travel, in one day, on the busiest avenue in the city: discrete quantitative.  
 Customer satisfaction with the service: excellent, good, regular, bad, or very bad: ordinal qualitative.  
 Blood type: nominal qualitative.  
 Email of Company X customers: nominal qualitative.  
 Time of professional experience of employees of Company X, in years: discrete quantitative.  
 Time of professional experience of employees of Company X, in seconds: discrete quantitative.  
 Professional experience of Company X employees: high experience, medium experience, or low experience: ordinal qualitative.  
 Gender of the interviewee: nominal qualitative.   
 HDI of the countries: continuous quantitative.  
 The inflation rate for the year: continuous quantitative.  

## Fix the Code

## 2.1  
```{r}
#"2 + 2"
2+2
```

## 2.2  
```{r}
#2 + "pi"
2+pi
```

## 2.3
```{r}
#1 = TRUE
x = TRUE
```

## 2.4
```{r}
#4 \ 3
4/3
```

##2.5
```{r}
#Help(cor.test)
help(cor.test)
```

## 2.6
```{r}
#???cor.test
?cor.test
```

## 2.7
```{r}
#help(if)
help("if")
```

## 2.8
```{r}
?help("cor.test")
help("cor.test")
?cor.test
?help
```

## 2.9
this is my sum
```{r}
# 1 + 2
1+2
```

## 2.10
```{r}
#numbers <- c(1 2 3 4)
numbers <- c(1, 2, 3, 4)
```

## 2.11
```{r}
#animals <- c("dog" "cat" "parrot" "bunny")
animals <- c("dog", "cat", "parrot", "bunny")
```

## 2.12
```{r}
#animals <- C("dog", "cat", "parrot", "bunny")
animals <- c("dog", "cat", "parrot", "bunny")
```

## 2.13
```{r}
#Head(mtcars)
head(mtcars)
```

## 2.14
```{r}
#str(Mtcars)
str(mtcars)
```

## 2.15
```{r}
#dim[mtcars]
dim(mtcars)
```

## 2.16
```{r}
#nomes(mtcars)
names(mtcars)
```

## 2.17
```{r}
#head(mtcars, x = 10)
head(mtcars, c = 10)
```

## 2.18

```{r}
#animals <- c("dog", "cat", "parrot", "bunny")
#numbers <- c(1,2,3,4,5,6,7,8,9,10)
#dat <- data.frame(animals,numbers)

animals <- c("dog", "cat", "parrot", "bunny")
numbers <- c(1,2,3,4)
dat <- data.frame(animals,numbers)
```

## 2.19
```{r}
dat <- c(1:100)
#save("my_data.RData", file = dat)

save(dat, file = ".RData")
```

## 2.20
````{r}
data("diamonds")
#ggplot(data = aes(x = cut)) +
#geom_bar()

library(ggplot2)
data("diamonds")
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```  
## 2.21
```{r}
data("diamonds")
#ggplot(data = diamonds) +
#geom_bar(mapping = aes(x = cut, y = cut))

library(ggplot2)
ggplot(data = diamonds) +
geom_bar(mapping = aes(y = cut))
```

## 2.22
```{r}
#theme_bw() +
#ggplot(data = diamonds, mapping = aes(x = cut))

a = ggplot(data = diamonds)+geom_bar(mapping = aes(x = cut))
a + theme_bw()
```n

ggplot(data = diamonds) +
geom_bar(mapping = aes(x = cut, fill = "red"))
```